---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r reading, cache=TRUE}
activity <- read.csv("data/activity.csv")
str(activity)
summary(activity)
```


## What is mean total number of steps taken per day?
First we aggregate the number of steps per day (you'll see a snapshot printed).
```{r aggregate and print the number of steps per day}
total_steps_perday <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
head(total_steps_perday)
```

And here you can see the how the user's activity varies and the associated mean and median.

```{r print a barplot, and calculate the mean and median}
library(ggplot2)
qplot(total_steps_perday$steps, xlab="Total steps per day", ylab="Frequency")
mean(total_steps_perday$steps, na.rm=TRUE)
median(total_steps_perday$steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
steps_by_interval <- aggregate(steps ~ interval, activity, mean)
plot(steps_by_interval$interval,steps_by_interval$steps,type="l",xlab= "Interval", ylab="Average Number of Steps", main="Average Number of Steps by Interval over time")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

````{r max number of steps}
max_interval <- steps_by_interval$interval[which.max(steps_by_interval$steps)]
max_interval

```

## Imputing missing values
The total number of missing values in the dataset (i.e. the total number of rows with NAs).
```{r number of missing values}
missing <- is.na(activity$steps)
table(missing)
```
Replacing the missing values with the interval's mean.

```{r replace missing with mean}
corr_activity <- activity
missing_boolean <- is.na(corr_activity$steps)
avg_interval <- tapply(corr_activity$steps, corr_activity$interval, mean, na.rm=TRUE)
corr_activity$steps[missing_boolean] <- avg_interval[as.character(corr_activity$interval[missing_boolean])]
```
Show corrected activity of the user per interval and the associated meand and median.

```{r print the corrected results}
corr_total_steps_perday <- aggregate(steps ~ date, data = corr_activity, FUN = sum, na.rm = TRUE)
head(corr_total_steps_perday)
qplot(corr_total_steps_perday$steps, xlab="Total steps per day with nas corrected", ylab="Frequency")
mean(corr_total_steps_perday$steps, na.rm=TRUE)
median(corr_total_steps_perday$steps, na.rm=TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r determine weekday or weekend}
library(lubridate)
library(dplyr)
library(plyr)
corr_activity <- corr_activity%>%mutate(typeofday= ifelse(weekdays(as.Date(corr_activity$date))=="zondag" | weekdays(as.Date(corr_activity$date))=="zaterdag", "Weekend", "Weekday"))
```

Calculate the mean for each interval.

```{r plot by type of day}
average_steps_bytypeofday <- aggregate(steps ~ interval + typeofday, data = corr_activity, FUN = mean, na.rm = TRUE)
head(average_steps_bytypeofday)
```

````{r plot by typeofday}
ggplot(corr_activity, aes(x =interval , y=steps, color=typeofday)) +
       geom_line() +
       labs(title = "Average steps by type of day", x = "Interval", y = "Average Steps") +
       facet_wrap(~ typeofday, ncol = 1, nrow=2)
```
